Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2018-07-26 ms.tex
brief updates

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby the action of self-gravity and radiative cooling alone produces galaxies with far too many stars. This problem has been addressed by employing various models of stellar feedback physics which are capable of generating self-regulating star formation in galaxies. Recently, it has become clear that the simplest models for stellar feedback, which account for supernovae (SNe) alone, are insufficient \citep{Smith2018} to fully regulate star formation and set galaxy properties consistent with observations (see \cite{NaabOstriker2017} for a recent review). Processes, such as stellar radiation, cosmic rays, or AGN, are often employed as additional sources of feedback. Modeling these processes in detail, however, is challenging. As a result, the relative importance of these additional physics is still uncertain.

@@

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Str{\"o}mgren approximation whereby the particles / cells within the Str{\"o}mgren radius around a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2016}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. Although the approximate radiation feedback models are capable of driving significant galactic outflows, it is unclear if they capture the full effects of stellar ionizing radiation. 
%The 
%It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds.

In this work we present a comparison of three simulations of the evolution of an isolated, low mass dwarf galaxy that account for detailed stellar feedback. Our fiducial model uses an adaptive ray-tracing radiative transfer method to follow stellar ionizing radiation. We test its effectiveness in regulating star formation and driving galactic winds by comparison to a simulation without ionizing radiation and a simulation with only localized radiation around each star. We discuss our methods in Section~\ref{sec:methods} and present our results in Section~\ref{sec:results}.

@@

We compare the resulting star formation rate (left) and gas mass properties (right) of our three simulations in Figure~\ref{fig:sfr_mass_evolution}. There is a clear, significant contrast between the runs with and without ionizing radiation. Stellar ionizing radiation leads to a factor of $\sim$ 5 reduction in the resulting SFR, as compared to the noRT run. Since the shortrad and fiducial simulations are so similar over the first $\sim$~100~Myr, it is clear that stellar radiation acts to significantly reduce the local star formation efficiency around young, massive stars. Radiation from these stars quickly ionizes and dissipates surrounding dense gas that would otherwise have gone to fuel a significant amount of additional star formation. We are unable to follow the long-term evolution of the noRT galaxy due to computational constraints from the large number of stars formed. Although radiative transfer is computationally expensive, this run is substantially more expensive than the fiducial and shortrad runs due to the significantly higher star formation rate.

>>>>

\section{Introduction} \label{sec:intro}
Historically, simulations of galaxy formation have suffered from the ``overcooling'' problem, whereby the action of self-gravity and radiative cooling alone produces galaxies with far too many stars. This problem has been addressed by employing various models of strong stellar feedback physics which are capable of generating self-regulating star formation in galaxies. \textbf{Energy injection from supernovae (SNe) has historically been used as the sole form of feedback. However, this is generally done heuristically as many simulations lack the ability to resolve the Sedov phase of individual SNe. But even with this strong feedback, recent work has argued for the need for pre-SN feedback, from stellar winds and/or stellar radiation, typically modelled simply as additional energy injection around newly formed stars (citations). Works that are capable of fulling resolving individual SN \citep{Hu2018,Smith2018} find that XXXXX Modelling these processes in detail, however, is challenging.

%Recently, it has become clear that the simplest models for stellar feedback, which account for supernovae (SNe) alone, are insufficient \citep{Smith2018} to fully regulate star formation and set galaxy properties consistent with observations (see \cite{NaabOstriker2017} for a recent review). Processes, such as stellar radiation, cosmic rays, or AGN, are often employed as additional sources of feedback. Modeling these processes in detail, however, is challenging. As a result, the relative importance of these additional physics is still uncertain.

@@

However, most works that employ stellar radiation feedback to account for these effects do so using various forms of sub-grid, approximate models to avoid the substantial additional cost of full radiative transfer. Many works use a Str{\"o}mgren approximation whereby the particles / cells within the Str{\"o}mgren radius around a radiating star are heated and kept ionized, with additional approximations made to account for overlapping ionized regions \citep[e.g.][]{HQM2011,Hu2016,Hu2017}. Other works employ some form of energy or momentum injection localized to the region immediately around a star particle \citep[e.g.][(\textit{need more})]{Agertz2013,Roskar2014,Ceverino2014,Forbes2016}. Although some of these approximate methods account for long-range effects of diffuse radiation \citep{HQM2012,Hopkins2018} most cases treat local radiation {\it only}, confined to energy or momentum injection in a limited physical region around a star particle. This is in part because the local destruction of dense clumps of gas around newly formed stars is commonly believed to be the dominant impact of stellar radiation feedback. The role of long-range stellar radiation, once ionizing photons break out of star forming regions, is not well characterized. Indeed it remains to be seen if modelling only the short-range effects of stellar radiation feedback is sufficient.

%Although the approximate radiation feedback models are capable of driving significant galactic outflows, it is unclear if they capture the full effects of stellar ionizing radiation. 
%The 
%It is unclear what, if any, effect the long-range ionization and heating processes play in regulating star formation and driving galactic winds.

In this work we use a detailed model for stellar feedback presented in \citep{Emerick2018} to study the role of radiation feedback in dwarf galaxy evolution. For the first time in a galaxy-scale simulation, we resolve the HII regions of single stars using an adaptive ray tracing radiative transfer method to follow the ionizing radaition from particles that represent individual stars. We focus on adressing two questions: 1) how important is radiation feedback in regulating star formation, and 2) are the long-range effects of radiation feedback important, or is the local destruction of dense gas the dominant effect. To investigate this we compare three simulations of the evolution of an isolated, low mass dwarf galaxy. Our fiducial model, containing full stellar radiation feedback, is compared against a run without ionziing radiation feedback, and a run with ionizing feedback limited to the local region around a given star. We dicuss our methods in Section~\ref{sec:methods} and present our results ion Section~\ref{sec:results}.

@@

We compare the resulting star formation rate (left) and gas mass properties (right) of our three simulations in Figure~\ref{fig:sfr_mass_evolution}. There is a clear, significant contrast between the runs with and without ionizing radiation. Stellar ionizing radiation leads to a factor of $\sim$ 5 reduction in the resulting SFR, as compared to the noRT run. Since the shortrad and fiducial simulations are so similar over the first $\sim$~100~Myr, it is clear that stellar radiation acts to significantly reduce the local star formation efficiency around young, massive stars. Radiation from these stars quickly ionizes and dissipates surrounding dense gas that would otherwise have gone to fuel a significant amount of additional star formation. We are unable to follow the long-term evolution of the noRT galaxy due to computational constraints from the large number of stars formed. Although radiative transfer is computationally expensive, this run is substantially more expensive than the fiducial and shortrad runs \textbf{due to a lower typical timestep from more shock-heated SN gas and additional work computing the stellar wind and optically thin radiation effects from more star particles}.
